### 2023-03-06 paper.tex
Add text and todos

\section{Introduction}
This paper will take a look at three popular language models utilizing
transformers in their architecture. It will also give a quick overview
of some applications and additions to the generic model.

TODO: explain why I chose these specific models to look at maybe?

\section{Training}
TODO: write about transfer learning in general, take from bert.
means big pretraining and then using that strong understanding
to fine tune for specific tasks

@@

@@

\subsection{Text-to-Text Transfer Transformer}
\begin{tikzpicture}[line width=0.3mm, align=center, font=\sffamily]
  % Nodes
  \matrix [matrix of nodes, row sep=2mm] (rightmat) {
    \node (translatein) [rounded corners=3pt, draw, fill=green!20] {"translate English to German: That is good."}; \\
    \node (colain)[rounded corners=3pt, draw, fill=red!20] {"cola sentence: The\\course is jumping well."}; \\
    \node (stsbin)[rounded corners=3pt, draw, fill=yellow!20] {"stsb sentence1: The rhino grazed\\on the grass. sentence2: A rhino\\is grazing in a field."}; \\
    \node (summaryin)[rounded corners=3pt, draw, fill=cyan!20] {"summarize: state authorities\\dispatched emergency crews tuesday to\\survey the damage after an onslaught\\of severe weather in mississippi..."}; \\
  };
  \node [right=of rightmat, rounded corners=3pt, draw, fill=gray!20] (t5) {\Huge T5};
  \matrix [matrix of nodes, right=of t5, row sep=2mm] {
    \node (translateout)[rounded corners=3pt, draw, fill=green!20] {"Das ist gut."}; \\
    \node (colaout)[rounded corners=3pt, draw, fill=red!20] {"not acceptable"}; \\
    \node (stsbout)[rounded corners=3pt, draw, fill=yellow!20]  {"3.8"}; \\
    \node (summaryout)[rounded corners=3pt, draw, fill=cyan!20] {"six people hospitalized after\\a storm in attala county."}; \\
  };

@@

  \draw [-] (translateout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
  \draw [-] (colaout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
  \draw [-] (stsbout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
  \draw [-] (summaryout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
\end{tikzpicture}

@@

\subsection{Toolformer}
Transformers can be taught to use different tools \cite{schick_toolformer_2023}. 
This can be achieved using additional fine tuning, as is the case with Toolformer
\cite{schick_toolformer_2023}, or by using few-shot methods telling the model 
about the tools available to it, as in Toolformer Zero \cite{markus_toolformer_2023}.

>>>>

\section{Introduction}
This paper will take a look at three popular language models utilizing
transformers in their architecture. It will also give a quick overview
of some applications and additions to the generic model. The aim is to
give the reader a general overview of the current landscape of different
transformer based language model architectures and some example applications.

The three different models are all similar in architecture being mostly
faithful to the original transformer architecture as presented in 
\cite{vaswani_attention_2017}. However, each of them have a different approach
to using the decoder and encoder stacks of the transformer. GPT uses only
the decoder, BERT only the encoder and T5 both encoder and decoder
\cite{radford_improving_nodate,devlin_bert_2019,raffel_exploring_2020}.

\section{Training}
All three of the models this paper discusses use transfer learning for 
training. The general idea is to first train a baseline model unsupervised 
using a large amount of unstructured text and then fine tuning the model
for each specific task after.

The objective of the unsupervised pre-training stage is to teach the model
general language understanding. The exact details between models varies but
in general, a massive number tokens are fed into to model during the course
of training. Each word corresponds to a single or a few tokens depending on 
the encoding used [citation needed, same as pricing] and the number of tokens
used in training can go up to the trillions \cite{raffel_exploring_2020}. 
The model is asked to predict a missing token based on context.

In the fine-tuning stage, the pre-trained model is used as a baseline. A new
model is trained using supervised learning with a task specific data set. 
The training can fine-tune all the parameters of the model, as in BERT 
\cite{devlin_bert_2019}, or an additional layer can be added and then
trained, as in GPT \cite{radford_improving_nodate}.

@@

TODO: talk about masking the attention layers somewhere, was talked about in
the t5 paper

TODO: talk about the exact differences between fine-tuning,
i think bert modifies the weights and gpt only adds an extra layer after

TODO: elmo for feature based?

@@

\subsection{Text-to-Text Transfer Transformer}
\begin{figure*}
  \centering
  \begin{tikzpicture}[line width=0.3mm, align=center, font=\sffamily]
    % Nodes
    \matrix [matrix of nodes, row sep=2mm] (rightmat) {
      \node (translatein) [rounded corners=3pt, draw, fill=green!20] {"translate English to German: That is good."}; \\
      \node (colain)[rounded corners=3pt, draw, fill=red!20] {"cola sentence: The\\course is jumping well."}; \\
      \node (stsbin)[rounded corners=3pt, draw, fill=yellow!20] {"stsb sentence1: The rhino grazed\\on the grass. sentence2: A rhino\\is grazing in a field."}; \\
      \node (summaryin)[rounded corners=3pt, draw, fill=cyan!20] {"summarize: state authorities\\dispatched emergency crews tuesday to\\survey the damage after an onslaught\\of severe weather in mississippi..."}; \\
    };
    \node [right=of rightmat, rounded corners=3pt, draw, fill=gray!20] (t5) {\Huge T5};
    \matrix [matrix of nodes, right=of t5, row sep=2mm] {
      \node (translateout)[rounded corners=3pt, draw, fill=green!20] {"Das ist gut."}; \\
      \node (colaout)[rounded corners=3pt, draw, fill=red!20] {"not acceptable"}; \\
      \node (stsbout)[rounded corners=3pt, draw, fill=yellow!20]  {"3.8"}; \\
      \node (summaryout)[rounded corners=3pt, draw, fill=cyan!20] {"six people hospitalized after\\a storm in attala county."}; \\
    };

@@

    \draw [-] (translateout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
    \draw [-] (colaout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
    \draw [-] (stsbout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
    \draw [-] (summaryout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
  \end{tikzpicture}
  \caption{Illustrating the types of inputs and outputs. Recreated from \cite{raffel_exploring_2020}.}
\end{figure*}

Text-to-Text Transfer Transformer, abbreviated as T5, is a language model modelling each
task as text in, text out. It uses both the encoder and decoder layers of the transformer.
\cite{raffel_exploring_2020}

The main differentiating factor of T5 is the way it chooses to model tasks. Instead of 
tailoring each output to the specific task at hand, every task takes in a text prefix
describing the task and outputting the results as text. For example, for STS-B classification
task the model is trained to out one of a set of floating point numbers 
$x \in \{ 1+0.2k \mid k \in \{ 0, 1, ..., 20\} \}$ as a text. Many other models just 
output a number directly as a sliding scale from 1 to 5.

@@

\subsection{Toolformer}
Transformers are able to produce very natural text but they struggle with seemingly
simple tasks such as simple mathematical calculations and factual information gathering.
To get around of these limitations, transformers can be taught to use external tools
to offload these tasks to applications better suited to them. \cite{schick_toolformer_2023} 
This can be achieved using additional fine tuning, as is the case with Toolformer
\cite{schick_toolformer_2023}, or by using few-shot methods telling the model 
about the tools available to it, as in Toolformer Zero \cite{markus_toolformer_2023}.

TODO: maybe an example here about asking chat gpt about a recent event leading to hallucination
and/or asking about a simple mathematical calculation leading to the wrong result
